---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# netpanel: An R Package that Contains Maximum Likelihood Estimation Routines for Panel Network Autocorrelation Models

<!-- badges: start -->

[![R-CMD-check](https://github.com/kevinCarson/netpanel/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/kevinCarson/netpanel/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `netpanel` package contains maximum likelihood estimation routines to fit panel network autocorrelation models (PNAMs) with time-varying social adjacency matrices. The 
package contains functionality for (1) fixed and random effects PNAMs, (2) dynamic PNAMs that include time lags of the outcome and networks, and (3) the estimation of network mixed 
effects models where, in addition to the lag of the outcome, the time-varying residuals are assumed to exhibit network autocorrelation. Importantly, the package allows for multiple 
network lags to be fitted simultaneously. Please see the list of below references for more information on panel network (spatial) autocorrelation models. For generality, a panel 
network autocorrelation model is a linear regression model that takes the following form:

$$Y_t = \lambda_1 W^{t}_{1}Y_t + \cdots + \lambda_k W^{t}_{k}Y_t + X_t\beta + v_t, \ \ t = 1,2,\cdots,T $$
where $Y_t$ is the outcome vector for $N$ actors at time $t$, $\lambda_k W^{t}_{k}Y_t$ is the network autocorrelation term for the $k^{th}$ social network at time $t$, $X_t$ are the set of exogenous time-varying (and invariant) regressors. Moreover, the broad
definition of the time-varying error term $v_t$ is:
$$v_t = \mu_N + e_t,\ \ t = 1,2,\cdots,T $$
where $\mu_N$ are time-invariant unit-specific unobserved factors thought to explain $Y_t$ and $e_t$ are the time- and unit-varying unobserved factors.

## Authors

**Kevin A. Carson**  
  - Author & Maintainer    
  - PhD Candidate at the [University of Arizona School of Sociology](https://sociology.arizona.edu/)  
  - Email: kacarson@arizona.edu  
  - Website: https://kevincarson.github.io/  

## Installation

You can install the developmental version of `netpanel` from [GitHub](https://github.com/) with:

``` r
remotes::install_github("kevinCarson/netpanel")
```

## The `netpanel` Package API

The `netpanel` package API primarily works through the wrapper function `mlpnam()`. The `mlpnam()`
function returns an S3 object of class `pnam` that stores the relevant results of the fitted PNAM. The full set of functions (including user-helper S3 methods) are:

  - `mlpnam()`: fits a panel network autocorrelation model to an empirical panel network dataset via Maximum Likelihood Estimation.
  - `netimpacts()`: computes the direct, indirect, and total network impacts for each included regressor from a fitted `pnam` S3 object.
  - `coef.pnam()`: extracts the vector of ML parameter estimates from a fitted `pnam` S3 object.
  - `vcov.pnam()`: extracts the asymptotic variance-covariance matrix from a fitted `pnam` S3 object.
  - `logLik.pnam()`: extracts the estimated log-likelihood from a fitted `pnam` S3 object.
  - `residuals.pnam()`: extracts the model residuals from a fitted `pnam` S3 object.
  - `fitted.pnam()`: extracts the linear and network regression fitted values from a fitted `pnam` S3 object.

## Fitting a Random Effects Panel Network Autocorrelation Model with `netpanel`

Importantly, the `netpanel` package includes a set of simulated pseudo data objects for users to start fitting panel network 
autocorrelation models. Below, we rely upon them to provide a brief example of how the `netpanel` package fits these types of 
models.

```{r processevents}
library(netpanel)
data("net1", package = "netpanel") #the first panel network
data("net2", package = "netpanel") #the second panel network
data("net3", package = "netpanel") #the third panel network
data("simulated.data", package = "netpanel") #the simulated variables 

# a random effects panel network autocorrelation model
re.pnam <- mlpnam(Y~x1+x2+x3, net.formula = ~ net1 + net2,
               data =  simulated.data, time = ~panel,
               actor = ~unit, model = "random")
```

```{r print, echo=TRUE}
summary(re.pnam) 
```
```{r computeimpacts}
impacts <- netimpacts(re.pnam) #extracting the network impacts
```

```{r print1, echo=TRUE}
print(impacts) 
```

We can extend the above model by fitting a mixed network random effects panel autocorrelation with `netpanel` that 
allows for the time-varying residuals to be autocorrelated across the network ties.

```{r mixedeffectsmodel}
# a random effects network mixed effects panel autocorrelation model
re.pnam.mixed <- mlpnam(Y~x1+x2+x3, net.formula = ~ net1 + net2,
               data =  simulated.data, time = ~panel,
               actor = ~unit, model = "random",
               errors = "autocorrelated",
               autocorrelated.network = net3)
```

```{r printmixed, echo=TRUE}
summary(re.pnam.mixed) 
```

## Questions, Comments, or Suggestions!

If you have any questions, comments, or suggestions please feel free to open an issue! 


